Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization

### 2023-10-30 manuscript.Rmd
Apply suggestions from maximilian
Co-authored-by: Maximilian-Stefan-Ernst <34346372+Maximilian-Stefan-Ernst@users.noreply.github.com>

This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].

@@

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study by investing resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974] or as recent psychological literature come to call it "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019].
That is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.

@@

The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
In the tradition of Carnap, in its direct use, it is called _confirmation as firmness_; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
In reality, researchers surely differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
The goal is therefore to reason about the space of possible measures researchers might apply.
However, since it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.

@@

The last statement sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
He stated that if it is impossible to prove that a hypothesis is false ($P(\neg{}E|\neg{}H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but Popper would disagree with attaching any form of believe in $H$ on the basis of evidence beyond failing to disprove it).

@@

One widely known factor is the contribution of statistical methods to theoretical risk.
Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative-hypothesis because then the nill-hypothesis is $\neg{}H$.
Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
Researchers who choose a smaller type-I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
However, this connection should not be overinterpreted for two reasons.
First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
Second, the research hypothesis rarely equals the statistical alternative hypothesis (most research hypothesis are more specific then "any value except zero").
In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers' belief.
One may take such theoretical properties as a first starting point to form a substantive belief but surely researchers ought to take other factors into consideration.
For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, this should decrease theorethical risk; after all they might find an association purely on account of the confounders.

@@

```{r other-measures-fig, cache=TRUE, warning=FALSE}
#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line). Measure taken from @sprengerBayesianPhilosophyScience2019, Table 1.3, p. 51."

@@

The argument for a confirmatory research agenda is that by increasing theoretical risk we increase expected epistemic value, i.e., moving to the right on the x-axis in Figure \ref{fig:posterior} increases posterior probability (on the y-axis).
However, if a hypothesis in a certain study has low theoretical risk, there is not much researchers can do about it.
However, studies do not only differ by how high the theoretical risk is but also by how certain the recipient is about the theoretical risk.
A study that has a very high theoretical risk (e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed,) but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extent of the benefit varies, of course, with the specifics of the study.
Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
We find it helpful to calculate some example because of the nonlinear nature of the evidence functions.

@@

1. We know with absolute certainty that reported results came about by p-hacking.
2. A hypothesis was picked to explain reported results after the fact [HARKing, @kerrHARKingHypothesizingResults1998].
3. We cannot exclude the possibility of p-hacking having led to the reported results.
4. Reported results were obtained by planned exploration.
5. Reported results were obtained by unplanned exploration.

In case 1, there is no theoretical risk ($P(\neg{}E|\neg{}H) = 0$).
If we know that the results will be engineered to support the hypothesis no matter what, there is no reason to collect data.
A prime example of this case is the $p_\text{pointless}$ metric [@husseyMethodStreamlinePhacking2021].
Case 2 has a similar problem.
After all, the hypothesis that it had to happen the way it did happen is irrefutable.
In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
Case 3, is different since here the theoretical risk is not necessarily low but simply uncertain (and perhaps best represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}).
In case 4, the theoretical risk is neither zero (unless the researcher plans to do run variations of analyses until a favourable outcome is obtained, then we have a particular instance case of 1) nor high (as this is the nature of exploratory approaches).
However, we can take advantage of computational reproducibility, use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
This favorable position leads us to recommend preregistration of exploratory studies.
Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard did researcher try to come up with favourable results).
Its low theoretical risk combined with high uncertainty about it, make it difficult to produce compelling evidence.

@@

Theoretical risk may conceptually be related to the conceptual framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
However, there are crucial differences between the two.
First, our perspective on theoretical risk is not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .999.
Second, severity considerations are mainly after the fact.
Severity, a measure with which we can rule out alternative explanations, can only be calculated after evidence was observed.
However, there also are communalities, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].

@@

Our emphasis on transparency aligns with other justifications of preregistration, especially those put forth by @lakensValuePreregistrationPsychological2019's, although based on quite different philosophical foundations.
Our goal is to contribute a rationale that more comprehensively captures the spectrum of exploration and confirmation in relation to preregistrations, post-hoc changes of preregistrations, and subjective evaluations of evidence.
We find it difficult to content ourselves with vague terms like "control" or "transparency" if they ultimately remain unconnected to how much researchers believe in a theory.
Within our framework, researchers have the ability to input their assumptions regarding the perspectives of other researchers and calculate the potential impact of their actions on their readership, whether these actions relate to study design, to the preregistration itself, or subsequent deviations from it.
We put subjective evaluations at the centre of our consideration; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence then others (they differ in their confirmation function).
We, therefore, hope to not only provide a rationale for preregistration for those who subscribe to a Bayesian philosophy of science but also a framework to navigate the complicated questions that arise in the practice of preregistration.

At the same time, approaching the evaluation of evidence using a Bayesian formalism is far from novel [e.g., @kuklaClinicalStatisticalTheory1990 @fiedlerWhatConstitutesStrong2017;@sprengerBayesianPhilosophyScience2019].
To our knowledge, it was not yet applied to the problem of preregistration.
However, @oberauerAddressingTheoryCrisis2019 made use of the formalism to model the relation between theory, hypothesis, and evidence.
In the context of this conceptualization, they discussed the usefulness of preregistration, though without applying the formalism there.
Most importantly, they are rather critical of the idea that preregistration has tangible benefits.
Instead, they prefer multiverse analyses but contend that those could be preregistered if one fancies it.
Their reasoning is based on two intuitions about what should *not* influence the evaluation of evidence: temporal order and the mental state of the originator.
For one, they disregard the temporal order a bit too hastily, as it is a long-standing issue in Bayesian philosophy of science known as the "problem of old evidence" [@chiharaProblemsBayesianConfirmation1987].
However, we agree that the information flow matters for preregistration, not the temporal order.
For the other, we argue that the mental state of the originator does matter.
Suppose there are $k = 1, 2, \ldots, K$ ways to analyze data.
Where each $k$ has a $P(E_k | \neg H) > 0$.
If they intend to try each way after another but happen to be "lucky" on the first try and stop, should we then apply $P(\neg E| \neg H) = P(\neg E_1 | \neg H)$ or $P(\neg E| \neg H) = \prod_{k=1}^K P(\neg E_k | \neg H)$?
We think the latter.
However, this "Defeatist" intuition is not universally warranted and depends on what we take $H$ to mean specifically [@kotzenMultipleStudiesEvidential2013].
Addressing, this problem might benefit from combining @oberauerAddressingTheoryCrisis2019's idea of updating on two nested levels (theory-hypothesis layered on top of hypothesis-evidence) with our approach to modelling uncertainty.

>>>>

This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].

@@

The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
Our rationale is as follows.
Researchers who decide to conduct a certain study are actually choosing a study to bet on.
They have to "place the bet" by conducting the study by investing resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in what studies to conduct.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974], that is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.
In particular, "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019], a concept highlighted in recent psychological literature, seems to adhere to the statistical relevancy condition.

@@

The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
In the tradition of Carnap, in its direct use, it is called _confirmation as firmness_; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
In reality, researchers surely differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
The goal is therefore to reason about the space of possible measures researchers might apply.
However, since any measure fulfilling the statistical relevancy condition increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.

@@

The last statement sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
He stated that if it is impossible to prove that a hypothesis is false ($P(\neg{}E|\neg{}H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but we of course differ from Popper on the central issue of induction in science).

@@

One widely known factor is the contribution of statistical methods to theoretical risk.
Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative-hypothesis because then the nill-hypothesis is $\neg{}H$.
Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
Researchers who choose a smaller type-I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
However, this connection should not be overinterpreted for two reasons.
First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
Second, the research hypothesis rarely equals the statistical alternative hypothesis (most research hypothesis are more specific than "any value except zero").
In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers' belief.
One may take such theoretical properties as a first starting point to form a substantive belief but surely researchers ought to take other factors into consideration.
For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, this should decrease theorethical risk; after all they might find an association purely on account of the confounders.

@@

```{r other-measures-fig, cache=TRUE, warning=FALSE}
#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line). Measures taken from @sprengerBayesianPhilosophyScience2019, Table 1.3, p. 51."

@@

The argument for a confirmatory research agenda is that by increasing theoretical risk we increase expected epistemic value, i.e., moving to the right on the x-axis in Figure \ref{fig:posterior} increases posterior probability (on the y-axis).
However, if a hypothesis in a certain study has low theoretical risk, there is not much researchers can do about it.
However, studies do not only differ by how high the theoretical risk is but also by how certain the recipient is about the theoretical risk.
A study that has a very high theoretical risk (e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed,) but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extent of the benefit varies, of course, with the specifics of the study.
Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
We find it helpful to calculate an example because of the nonlinear nature of the evidence functions.

@@

1. We know with absolute certainty that researchers will revert to p-hacking to create evidence that is favorable for the theory.
2. A hypothesis was picked to explain reported results after the fact [HARKing, @kerrHARKingHypothesizingResults1998].
3. We cannot exclude the possibility of p-hacking having led to the reported results.
4. Reported results were obtained by planned exploration.
5. Reported results were obtained by unplanned exploration.

In case 1, there is no theoretical risk ($P(\neg{}E|\neg{}H) = 0$).
If we know that the results will be engineered to support the hypothesis no matter what, there is no reason to collect data.
A prime example of this case is the $p_\text{pointless}$ metric [@husseyMethodStreamlinePhacking2021].
Case 2 has a similar problem.
After all, the hypothesis that it had to happen the way it did happen is irrefutable.
In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
Case 3 is different since here the theoretical risk is not necessarily low but simply uncertain (and perhaps best represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}).
In case 4, the theoretical risk is neither zero (unless the researcher plans to do run variations of analyses until a favourable outcome is obtained, then we have a particular instance case of 1) nor high (as this is the nature of exploratory approaches).
However, we can take advantage of computational reproducibility, use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
This favorable position leads us to recommend preregistration of exploratory studies.
Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard did researcher try to come up with favourable results).
Its low *and uncertain* theoretical risk make it difficult to produce compelling evidence.

@@

Theoretical risk may conceptually be related to the framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
However, there are crucial differences between the two.
First, our perspective on theoretical risk is not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .999.
Second, severity considerations are mainly after the fact.
Severity, a measure with which we can rule out alternative explanations, can only be calculated after evidence was observed.
However, there also are communalities, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].

@@

Our emphasis on transparency aligns with other justifications of preregistration, especially those put forth by @lakensValuePreregistrationPsychological2019's, although based on quite different philosophical foundations.
Our goal is to contribute a rationale that more comprehensively captures the spectrum of exploration and confirmation in relation to preregistrations, post-hoc changes of preregistrations, and subjective evaluations of evidence.
We find it difficult to content ourselves with vague terms like "control" or "transparency" if they ultimately remain unconnected to how much researchers believe in a theory.
Within our framework, researchers have the ability to input their assumptions regarding the perspectives of other researchers and calculate the potential impact of their actions on their readership, whether these actions relate to study design, to the preregistration itself, or subsequent deviations from it.
We put subjective evaluations at the centre of our considerations; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence than others (they differ in their confirmation function).
We, therefore, hope to not only provide a rationale for preregistration for those who subscribe to a Bayesian philosophy of science but also a framework to navigate the complicated questions that arise in the practice of preregistration.

At the same time, approaching the evaluation of evidence using a Bayesian formalism is far from novel [e.g., @kuklaClinicalStatisticalTheory1990 @fiedlerWhatConstitutesStrong2017;@sprengerBayesianPhilosophyScience2019].
To our knowledge, it was not yet applied to the problem of preregistration.
However, @oberauerAddressingTheoryCrisis2019 made use of the formalism to model the relation between theory, hypothesis, and evidence.
In the context of this conceptualization, they discussed the usefulness of preregistration, though without applying the formalism there.
Most importantly, they are rather critical of the idea that preregistration has tangible benefits.
Instead, they prefer multiverse analyses but contend that those could be preregistered if one fancies it.
Their reasoning is based on two intuitions about what should *not* influence the evaluation of evidence: temporal order and the mental state of the originator.
In our opinion, they disregard the temporal order a bit too hastily, as it is a long-standing issue in Bayesian philosophy of science known as the "problem of old evidence" [@chiharaProblemsBayesianConfirmation1987].
However, we agree that not the temporal order is decisive but if the researchers incorporated the information into the hypothesis the evidence is supposed to confirm.
For the other, we argue that the mental state of the originator does matter.
Suppose there are $k = 1, 2, \ldots, K$ ways to analyze data.
Where each $k$ has a $P(E_k | \neg H) > 0$.
If they intend to try each way after another but happen to be "lucky" on the first try and stop, should we then apply $P(\neg E| \neg H) = P(\neg E_1 | \neg H)$ or $P(\neg E| \neg H) = P(\neg E_1 \land \ldots \land \neg E_k | \neg H)$?
We think the latter.
However, this "Defeatist" intuition is not universally warranted and depends on what we take $H$ to mean specifically [@kotzenMultipleStudiesEvidential2013].
Addressing, this problem might benefit from combining @oberauerAddressingTheoryCrisis2019's idea of updating on two nested levels (theory-hypothesis layered on top of hypothesis-evidence) with our approach to modelling uncertainty.
